### 2020-04-07 submission/manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  11pt,
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1.0in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font

@@

\usepackage{pdfpages}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{cslreferences}%
  {}%
  {\par}

@@

\hypertarget{ten-simple-rules-for-running-a-code-club-for-increasing-computational-skills-among-biologists}{%
\section{Ten Simple Rules for running a Code Club for increasing
computational skills among
biologists}\label{ten-simple-rules-for-running-a-code-club-for-increasing-computational-skills-among-biologists}}

@@

\emph{Author names here, order TBD}

@@

@@

\hypertarget{ten-simple-rules}{%
\subsubsection{Ten Simple Rules}\label{ten-simple-rules}}

\newpage
\linenumbers

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

For most biologists, the ability to generate data has outpaced the
ability to analyze those data. High throughput data comes to us from DNA
and RNA sequencing, flow cytometry, metabolomics, molecular screens, and
more. Although some accept the approach of compartmentalizing the
ability to generate data and the ability analyze data, we have found
scientists feel empowered when they can both ask and answer their own
biological questions. Yet, the standard undergraduate and graduate
training in the biological sciences is insufficient to develop the data
analysis skills needed to analyze one's data. In our experience
performing microbiome research, it is more common to find exceptional
bench scientists who are inexperienced at analyzing large datasets than
to find the reverse. Of course this raises a challenge: How do we train
bench scientists to analyze datasets to answer biological questions?

The ever-growing ability to generate data and constant feeling of
helplessness in analyzing it is analogous to the struggles we also face
with engaging the voluminous scientific literature. Within our research
group we have leveraged this similarity to address the challenges of
training bench scientists to engage in reproducible data analysis
practices. Our group and others hold a regular Journal Club to stay
current on the literature and model to junior scientists how to engage
that literature {[}1{]}. Over the past 4 years we have experimented with
using a Code Club to introduce new software packages, solve specific
problems, and evaluate each other's code.

Initially, the Code Club was created as a way to review code that was
being used in different projects. We would project someone's script onto
a screen and go line by line through the code with each person in the
room stating the logic behind the code. This approach emphasized the
importance of readability and it gave beginners the opportunity to see
the real, ``smelly'' code of more experienced people in the group.
Unfortunately, the format only allowed us to review a small fraction of
a project's code base and it became difficult to integrate the
programmer's logic across their full project. We quickly realized that
although beginners could contribute to the discussion, the more senior
group members dominated conversation. This frequently created an
adversarial environment, which led to the presenter not offering to
present again. From this experience we learned that it is critical for
the presenter to clearly articulate their goals for the Code Club and
that the participants needed to engage with the code in that context.

Our Code Club sessions are each an hour long and are held roughly every
other week as part of a weekly two hour lab meeting. Presenters
voluntarily sign up to lead a session. The current format of our Code
Club generally takes one of two approaches. The first is a more humane
version of a group code review. The presenter will clearly state the
problem they would like to solve, break the participants into smaller
groups, and then ask each group to solve the problem. For example,
someone may have an R script where a chunk of code is repeated
throughout the script. The challenge for the session would be to convert
the chunk into a function and call it throughout the script to make it
``DRY'' (i.e.~Don't Repeat Yourself){[}2{]}. The presenter leaves the
session with several partial or working solutions to their problem and
everyone relearns the importance of writing DRY code. The second is a
tutorial approach. The presenter introduces a new package or way of
doing things. They then give the group an activity that forces them to
replace their current approach with a new approach. For example, one of
the more memorable activities was giving participants raw data and a
finished plot and asking two pairs of participants to generate the plot
using base R and asking the other two pairs to generate the plot with
ggplot. The stipulation was that base R users had to build the plots in
ggplot and ggplot users had to build the plot in base R. Each Code Club
ends with an opportunity to report back to the presenter and larger
group. Our Code Clubs typically have 7 to 10 participants. Because of
the ``think-pair-share'' approach that is baked into each session it
could likely be scaled to larger groups {[}3{]}.

We continue to experiment with approaches to running Code Club. Although
any individual Code Club session may be be more experimental than
others, on the whole they have been a critical tool to providing the
much needed training to use reproducible practices for analyzing data.
Examples of successful code club topics are described in Table 1. These
Ten Rules summarize what we have learned as presenters and participants.

\hypertarget{rule-1-reciprocate-respect}{%
\subsection{Rule 1: Reciprocate
respect}\label{rule-1-reciprocate-respect}}

It is critical that the presenter and participants respect each other
and that there be an individual (e.g.~the lab director) who enforces a
code of conduct. Everyone in the session needs to have the humility to
acknowledge that they still have more to learn about any topic. We have
found that many problems are avoided when the presenter takes charge of
the session by having a clear lesson plan, thoughtfully creating groups,
and giving encouragement. Similarly, participants foster a positive
environment by remembering that the task is not a competition, focusing
on the presenter's goals, allowing their partner to contribute, asking
clarifying questions when appropriate, and avoiding distractions
(e.g.~email, social media). Learning to program is challenging and too
often it can lead to a toxic environment. All parties in a Code Club are
responsible for preventing this toxicity.

\hypertarget{rule-2-set-specific-goals}{%
\subsection{Rule 2: Set specific
goals}\label{rule-2-set-specific-goals}}

In our early Code Clubs we noticed that if the presenter did not clearly
state their goals for the session, it would often led to frustration for
the presenter and participants. If the presenter shared their own code,
did they want us to focus on their coding style or do they want help
with incorporating a new package into their workflow? Participants will
always notice or ask about things that are not the focus of the
exercise. Be specific about the goal and bring them back to the task if
they are distracted. Where possible, creating a simplified scenario
(i.e.~minimal, reproducible example {[}4{]}) can be helpful to remove
distractions. The presenter should make sure that the simplified example
works and behaves the way they expect before coming to the Code Club.
Beyond the content, participants should know what they can be expected
to have completed by the end of the session.

\hypertarget{rule-3-keep-it-simple}{%
\subsection{Rule 3: Keep it simple}\label{rule-3-keep-it-simple}}

Our Code Club needs to fit within an hour time slot. The presenter
should anticipate enough time for an introduction and brief instruction,
time for participants to engage the material, and time for everyone to
report back. A typical schedule for Code Review is 10 minutes of
introduction and instruction, 30 minutes of paired programming, 5
minutes to get groups to wrap up, and 10 minutes to report back to the
group. Remember that a task that you can accomplish in 10 minutes may
take a more junior member of your group 30 minutes. The goal of the Code
Club should be to help that junior member to learn the topic and to
reinforce the topic for the more senior person and help them see the
concept in a new context. One possibility would be to create ``stretch
goals'' for those that complete the task more quickly than everyone
else. We once had a presenter try to teach the group some basic Julia
syntax. By the time people installed the interpreter, the time was up.
Alternatively, a session where participants were asked to make code DRY
gave them multiple stopping points where they could feel successful
(e.g.~commenting code, creating function, implementing function,
refactoring function).

\hypertarget{rule-4-give-participants-time-to-prepare}{%
\subsection{Rule 4: Give participants time to
prepare}\label{rule-4-give-participants-time-to-prepare}}

Similar to a Journal Club, the presenter should give participants a few
days to prepare for the Code Club. Be sure to tell the participants what
you will be covering. Provide the them with instructions on how to
install dependencies, download data, and get the initial code. Make sure
that the materials work together as intended. Considering the compressed
schedule, asking participants to download materials beforehand is
helpful. We have found that using separate GitHub repositories for each
session can be a helpful way to get information to participants. Of
course, using git and GitHub to engage in collaborative coding should
probably be its own Code Club before using the tools more broadly with
the group. An easier entry point is to have groups post their code to a
Slack channel that we use to announce materials for our meetings.

\hypertarget{rule-5-prepare-in-advance-to-maximize-your-participation}{%
\subsection{Rule 5: Prepare in advance to maximize your
participation}\label{rule-5-prepare-in-advance-to-maximize-your-participation}}

It is the rare scientist that can fully participate in a Journal Club
discussion about a paper that they have not read. Coming to Code Club
without installing a dependency is like going to Journal Club and asking
a simple question about the paper. It shows a lack of respect and
preparation. Following Rule 4, review the code in advance, download the
datasets, install the necessary packages, and perhaps read up on the
topic. If your Code Club is reading a paper or chapters in a data
science book be sure to read them before the session (e.g.~{[}5--7{]}).
Consider how you could use the material for your own work.

\hypertarget{rule-6-dont-give-your-participants-busy-work}{%
\subsection{Rule 6: Don't give your participants busy
work}\label{rule-6-dont-give-your-participants-busy-work}}

Participants want to learn something because it will be useful to them
or because they think it will help the presenter to improve their code.
Presenters should do their best to satisfy those desires. It would not
make much sense to hold a Code Club on downloading stock market data if
the group was not interested in those data. At the same time,
participants should not be asked to help the presenter improve their
code if the presenter has no intention of incorporating the suggestions.
It would be useful to keep a list of packages or tasks that members of
the group want to learn about. That way if a presenter is having a hard
time coming up with ideas, they can take something from the list and
present it.

\hypertarget{rule-7-include-all-levels-of-participants}{%
\subsection{Rule 7: Include all levels of
participants}\label{rule-7-include-all-levels-of-participants}}

A significant challenge to presenting at Code Club is to pick topics and
activities that appeal to a critical mass of the participants. For
example, if there has recently been a large turn over in group
membership, it can be useful to go back to the basics and cover commonly
used tools. We have identified several strategies to overcome the
challenges of having varying skill levels. As an alternative to letting
participants form their own pairs, the presenter can be purposeful in
how they form pairs. Depending on the goals of the session, forming
pairs between participants that have similar or different skills can be
effective. If pairs are formed where partners have similar skill levels,
then the presenter needs to design activities that are appropriate to
different levels of participants. We have found that a good activity for
beginners is to have them comment code, which forces them to dissect a
code chunk and understand what each line does. An advantage of forming
pairs with disparate skill levels is that all of the pairs can be given
the same task leading to a diversity of methods to achieve the same
result. This approach to pairing also helps graft new members into the
group that have emerging programming skills. Regardless of how partners
are selected, asking the pairs to identify a navigator and driver is
helpful {[}8{]}. The driver types at the computer while the navigator
tells them what to type. Midway through the activity the presenter can
tell the partners to switch roles. Intentionally forming pairs can also
be a mechanism to engineer interactions by breaking up cliques, avoiding
potentially toxic combinations of members, or pairing reliable role
models with new members of the group.

\hypertarget{rule-8-participate}{%
\subsection{Rule 8: Participate}\label{rule-8-participate}}

Having an open laptop on the table and permission to use it can feel
like an invitation to get distracted by other work, emails, and browsing
the internet. Fight that urge and focus on the presenter's goals. Allow
your partner to contribute. Speak up for yourself and force your partner
to let you contribute. Even if the material seems too advanced for you,
programming languages like R and Python are generally expressive and
should allow you to engage with the logic, even if the syntax is too
hard. Frankly, understanding the logic of when to use one modeling
approach over another is more important than how to use the modeling
approach. If you understand the ``why'', the ``how'' will generally
quickly follow. If you are a more experienced participant, give your
feedback and coding suggestions at a level that seeks to inform
everyone. Regardless of your skill level, your partner and the presenter
are putting themselves at risk by revealing what they do or do not know.
Encourage them and show your gratitude for helping you learn something
new.

\hypertarget{rule-9-let-the-material-change-you}{%
\subsection{Rule 9: Let the material change
you}\label{rule-9-let-the-material-change-you}}

Part of the humility required to participate in a Code Club is both
acknowledging your incomplete training and that learning something new
is possible. Assume that the presenter has followed the other rules and
that they think you should be using what they presented. After the Code
Club, try to incorporate that material into your new code or by
refactoring your old code. By practicing the material in a different
context you will learn the material better. You may identify concepts
that you are unsure of, which gives you another opportunity to ask the
presenter for help.

\hypertarget{rule-10-experiment}{%
\subsection{Rule 10: Experiment!}\label{rule-10-experiment}}

The selection of content and structure for each Code Club is determined
by the presenter making the process democratic and distributed. If
someone thinks something is worth learning and wants to teach it, they
have that power. If they want to experiment with a different format,
they are free to try it out. Members of the Code Club need to feel like
they have the power to shape the direction of the group. If members are
following these Rules, they will naturally reflect on the skills and
interests of the other members in the group. For example, there is
always turn over in a research group and thus it is important to make
sure that there are opportunities to go back to things people consider
``simple'' to teach the content to new people or provide a refresher to
people that perhaps have been at the lab bench more than at the computer
recently. The group should also feel free to experiment with the format.
If we try a different format, then it is useful to give everyone an
opportunity to debrief on what they did or did not like with the new
approach.

\hypertarget{conclusion}{%
\subsection{Conclusion}\label{conclusion}}

The most important rules are the first and last. Members of the Code
Club need to feel comfortable with other members of the group and
sufficiently empowered to try something new or to ask for help. Aside
from growing in our programming skills, we have noticed two other
benefits that point to the efforts of creating a positive culture.
First, we have intentionally interviewed postdoc candidates on the days
we have held Code Club. Although it is important to make it clear that
they are not being assessed on their coding skills, it is useful to see
how the candidate interacts with other members of the research group in
a relatively low stakes environment. At the same time, the candidate can
learn about the culture of the research group. Second, members of other
research groups where they are isolated in their desire to grow in these
skills have integrated themselves into our Code Club. This speaks to the
broader need for Code Club and the likelihood that it would work when
expanded to a larger group of individuals that are not necessarily in
the same research group. Ultimately, our Code Club has helped
researchers feel more comfortable seeking out their colleagues outside
of Code Club sessions for help in improving their code.

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

All co-authors have been participants in the Code Clubs run as part of
PDS's lab meetings. They have had a critical role in shaping the
evolution of the sessions. The order of co-authors was determined
by\ldots{}

@@

\hypertarget{references}{%
\subsection{References}\label{references}}

\hypertarget{refs}{}
\begin{cslreferences}
\leavevmode\hypertarget{ref-Lonsdale2016}{}%
1. Lonsdale A, Penington JS, Rice T, Walker M, Dashnow H. Ten simple
rules for a bioinformatics journal club. Lewitter F, editor. PLOS
Computational Biology. 2016;12: e1004526.
doi:\href{https://doi.org/10.1371/journal.pcbi.1004526}{10.1371/journal.pcbi.1004526}

\leavevmode\hypertarget{ref-Wilson2014}{}%
2. Wilson G, Aruliah DA, Brown CT, Hong NPC, Davis M, Guy RT, et al.
Best practices for scientific computing. Eisen JA, editor. PLoS Biology.
2014;12: e1001745.
doi:\href{https://doi.org/10.1371/journal.pbio.1001745}{10.1371/journal.pbio.1001745}

\leavevmode\hypertarget{ref-Kothiyal2014}{}%
3. Kothiyal A, Murthy S, Iyer S. Think-pair-share in a large CS1 class.
Proceedings of the 2014 conference on innovation \& technology in
computer science education - ITiCSE 14. ACM Press; 2014.
doi:\href{https://doi.org/10.1145/2591708.2591739}{10.1145/2591708.2591739}

\leavevmode\hypertarget{ref-StackOverflow_MREE}{}%
4. Stack Overflow. How to create a minimal, reproducible example. 2020.
Available:
\url{https://stackoverflow.com/help/minimal-reproducible-example}

\leavevmode\hypertarget{ref-Wickham_2016}{}%
5. Wickham H, Grolemund G. O'Reilly Media; 2016. Available:
\url{https://r4ds.had.co.nz}

\leavevmode\hypertarget{ref-Schloss_2018}{}%
6. Schloss PD. The riffomonas reproducible research tutorial series.
Journal of Open Source Education. 2018;1: 13.
doi:\href{https://doi.org/10.21105/jose.00013}{10.21105/jose.00013}

\leavevmode\hypertarget{ref-Noble2009}{}%
7. Noble WS. A quick guide to organizing computational biology projects.
Lewitter F, editor. PLoS Computational Biology. 2009;5: e1000424.
doi:\href{https://doi.org/10.1371/journal.pcbi.1000424}{10.1371/journal.pcbi.1000424}

\leavevmode\hypertarget{ref-Bockeler_2020}{}%
8. Böckeler B, Siessegger N. On pair programming. 2020. Available:
\url{https://martinfowler.com/articles/on-pair-programming.html}
\end{cslreferences}

>>>>

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1.0in]{geometry}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font

@@

\usepackage{pdfpages}

@@

\section{Ten Simple Rules to increase computational skills among
biologists with Code
Clubs}\label{ten-simple-rules-to-increase-computational-skills-among-biologists-with-code-clubs}

@@

Ada K. Hagan\({^2}\), Nick Lesniak, \emph{Author names here, order TBD}

@@

\(2\) Alliance SciComm \& Consulting, Linden, MI 48418

@@

\subsubsection{Ten Simple Rules}\label{ten-simple-rules}

@@

\linenumbers

\subsection{Introduction}\label{introduction}

For most biologists, the ability to generate data has outpaced the
ability to analyze those data. High throughput data comes to us from DNA
and RNA sequencing, flow cytometry, metabolomics, molecular screens, and
more. Although some accept the approach of compartmentalizing data
generation and data analysis, we have found scientists feel empowered
when they can both ask and answer their own biological questions. Yet,
the standard undergraduate and graduate training in the biological
sciences is insufficient for developing these advanced data analysis
skills. In our experience performing microbiome research, it is more
common to find exceptional bench scientists who are inexperienced at
analyzing large data sets than to find the reverse. Of course this
raises a challenge: How do we train bench scientists that analyze data
sets to answer biological questions?

The ever-growing ability to generate data and constant feeling of
helplessness in analyzing it is analogous to the struggles we also face
with engaging the voluminous scientific literature. A common strategy to
keeping up with the literature are journal clubs, which involve group
discussion of a pre-selected paper. These range from informal
discussions to PowerPoint presentations and course credit. In addition
to staying current on the literature, journal clubs help strengthen
skills in critical thinking, communication, and literature review
{[}1{]}. Our research group has leveraged the similarity between the
overwhelming natures of scientific literature and data analysis to
address the challenge of teaching reproducible data analysis practices
to bench scientists. Over the past 4 years we have experimented with a
Code Club model to improve data analysis skills in a community
environment.

Our Code Club sessions are an hour long and alternate with Journal Club
as the second part of weekly two hour lab meetings. Presenters volunteer
to lead a session with the expectation that they alternate leading
Journal and Code Clubs. Initially, the Code Club was used to review code
from trainee projects. Instead of a presentation, the presenter would
project their code onto a screen and the participants would go line by
line through the code, stating the logic behind each line. This approach
emphasized the importance of code readability and gave beginners the
opportunity to see the real-life, messy code of more experienced peers.
Unfortunately, the format only allowed us to review a fraction of a
project's code base, making it difficult to integrate the programmer's
logic across their full project. A major issue with this model was that
sometimes beginners couldn't contribute to improving the code, and even
when they could, more senior group members would eventually dominate the
discussion. This led to a lack of participation by beginners, who would
mentally check out to focus on other activities, and frequently resulted
in an adversarial environment between the presenter and senior group
members. As a result, presenters rarely offered to present Code Club
again.

From these experiences, we began a collective conversation to improve
our Code Club model and have evolved it into two approaches. The first
is a more constructive version of a group code review. The presenter
clearly states the problem they want to solve, breaks the participants
into smaller groups, and then asks each group to solve the problem, or a
portion of it. For example, someone may have an R script with a
repeating chunk of code. The challenge for the session would be to
convert the code chunk into a function to be called throughout the
script to make it ``DRY'' (i.e.~Don't Repeat Yourself){[}2{]}. The
presenter leaves the session with several partial or working solutions
to their problem and the importance of writing DRY code is reinforced.
The second approach is a tutorial. The presenter introduces a new
package or technique and assigns an activity to practice the new
approach. For example, a more memorable activity was giving participants
raw data and a finished plot. Paired participants were tasked with
generating the plot using either base R or ggplot, but base R users had
to use ggplot and vice versa. In either approach, the Code Club ends
with a report back to the presenter and larger group describing the
approach each pair took. Our Code Clubs typically have 7 to 10
participants, but the inherent ``think-pair-share'' strategy should
allow it to be scaled to larger groups {[}3{]}.

We continue to experiment with approaches to running Code Club, but have
learned that it is critical for the presenter to clearly articulate
their goals and facilitate participant engagement. Although any
individual Code Club session may be more experimental than another, on
the whole they are a critical tool to train bench scientists in
reproducible data analysis practices. We have provided some examples of
successful code club topics in Table 1 and summarized the results of our
own experimentation as Code Club presenters and participants into Ten
Rules. The first 3 rules apply to all participants at every Code Club,
while Rules 4 through 8 are targeted to Code Club presenters and the
last two for participants.

\subsection{Rule 1: Reciprocate
respect}\label{rule-1-reciprocate-respect}

It is critical that the presenter and participants respect each other
and that a designated individual (e.g.~the lab director) enforces a code
of conduct (see X for an example). Each member of the Code Club must
have the humility to acknowledge that they have more to learn about any
given topic. We have found that many problems are avoided when the
presenter takes charge of the session with a clear lesson plan,
thoughtful group creation, and constructive encouragement. Similarly,
participants foster a positive environment by remembering that the task
is not a competition, focusing on the presenter's goals, allowing their
partner to contribute, asking clarifying questions when appropriate, and
avoiding distractions (e.g.~email, social media). Learning to program is
challenging and too often results in a toxic environment of nitpicking
and nonconstructive criticism. All parties in a Code Club are
responsible for preventing this toxicity by demonstrating respect for
themselves and their colleagues.

\subsection{Rule 2: Let the material change
you}\label{rule-2-let-the-material-change-you}

Part of the humility required to participate in a Code Club is
acknowledging that your training is incomplete and that it is possible
for you to learn something new. For participants, assume that the
presenter has a plan and follow their presented approach. After the Code
Club, try to incorporate that material into new code or refactor old
code. By practicing the material in a different context you will learn
the material better. For presenters, incorporate suggested changes into
your code. Either party may identify concepts that they are unsure of,
presenting opportunities for further conversation and learning.

\subsection{Rule 3: Experiment!}\label{rule-3-experiment}

The selection of content and structure for each Code Club is democratic
and distributed. If someone thinks a technique is worth learning and
wants to teach it, they have that power. If they want to experiment with
a different format, they are free to try it out. Members of the Code
Club need to feel like they have the power to shape the direction of the
group. If members are following these Rules, they will naturally reflect
on the skills and interests of the other members in the group. For
example, there is always turnover in a research group, making it
important to revisit basic concepts to teach it to new people and
provide a refresher to others. The group should also feel free to
experiment with the format and incorporate group feedback by ending with
a debrief discussing the pros and cons of new formats.

\subsection{Rule 4: Set specific goals}\label{rule-4-set-specific-goals}

In our early Code Clubs we noticed that if the presenter did not clearly
state their goals for the session, it often led to frustration for both
the presenter and participants. If the presenter shared their own code,
did they want participants to focus on their coding style or did they
want help incorporating a new package into their workflow? Participants
will always notice or ask about code concepts that are not the focus of
the exercise; a presenter with a specific goal can bring tangential
conversations back to the planned task. Where possible, presenters
should create a simplified scenario (i.e.~minimal, reproducible example
{[}4{]}), which can be helpful in focusing the participants. The
presenter should verify ahead of time that the simplified example works
and behaves the way they expect. Beyond the content, clear goals for
participant activities will help both parties stay on task and avoid
frustration. For more advanced learners, the presenter can create
stretch goals or give an activity with multiple stopping points where
participants would feel successful (e.g.~commenting code, creating
function, implementing function, refactoring function). Accomplishing
specific goals is more likely to result in a positive outcome for
presenters and participants.

\subsection{Rule 5: Keep it simple}\label{rule-5-keep-it-simple}

Our Code Club needs to fit within an hour time slot. When considering
their Code Club activity, the presenter should plan for an introduction
and brief instruction, time for participants to engage the material, and
time for everyone to report back within that hour. A typical schedule
for Code Club is 10 minutes of introduction and instruction, 30 minutes
of paired programming, 5 minutes to get groups to wrap up, and 10
minutes to report back to the group. Remember that learners may need up
to three times as long to complete a task that is simple for the
presenter, so Code Club is best kept simple. We once had a presenter try
to teach basic Julia syntax, but time was up by the time participants
installed the interpreter. Some tips to help keep it simple are to limit
the presented code chunks to less than 50 or, conversely, consider the
number of lines that might be required to accomplish a solution.

\subsection{Rule 6: Give participants time to
prepare}\label{rule-6-give-participants-time-to-prepare}

Similar to a Journal Club, the presenter should give participants a few
days (ideally a week) to prepare for the Code Club. Considering the
compressed schedule described in Rule 5, asking participants to download
materials beforehand is helpful to ensure a quick start. The presenter
should provide the participants with instructions on how to install
dependencies, download data, and get the initial code. This might also
uncover weak points in the presenter's plan and enable them to ensure
that the materials work as intended before the Code Club. We have found
that using GitHub repositories for each Code Club can help make
information, scripts, and data easily available to participants. Perhaps
a first Code Club could be using git and GitHub to engage in
collaborative coding. A lower barrier entry point is posting their code,
data, and information in a lab meeting-dedicated Slack channel or
sharing via email. Whatever method is used, the presenter should be sure
to communicate the topic and necessary materials with the participants
ahead of time.

\subsection{Rule 7: Don't give participants busy
work}\label{rule-7-dont-give-participants-busy-work}

Participants want to learn topics that will either be useful to them or
help their colleague (i.e., the presenter). Presenters should do their
best to satisfy those motivations, whether it is through the relevancy
of the concept or the data. It does not make sense to present a Code
Club on downloading stock market data if it is not useful or interesting
to the group. Similarly, participants should not be tasked with
improving the presenter's code if the presenter has no intention of
incorporating the suggestions. A list of packages or tasks that group
members are interested in could help an uninspired presenter choose a
topic that will make a rewarding Code Club.

\subsection{Rule 8: Include all levels of
participants}\label{rule-8-include-all-levels-of-participants}

As suggested by Rule 7, a significant challenge to presenting at Code
Club is selecting topics and activities that appeal to a critical mass
of the participants. This is particularly difficult if participants have
a wide range of coding experience, such as follows a turnover in group
membership. In these circumstances it can be useful to cover commonly
used tools but that might result in disinterest of more experienced
participants. We have identified several strategies to overcome the
challenges presented by participants at varying skill levels. Instead of
letting participants form their own pairs, the presenter can select
pairs of participants with either similar or differing skill levels,
depending on their goals. Partnerships between those with similar skill
levels requires the presenter to design appropriate activites each skill
level. We have found that commenting code is a good skill for beginners
since it forces them to dissect and understand a code chunk
line-by-line. An advantage of partnerships with disparate skill levels
is assigning a single task, which provides the presenter with a diverse
range of methods that achieve the same result. This approach to pairing
also helps to graft new members into the group that have emerging
programming skills. Regardless of how partners are selected, consider
asking the pairs to identify a navigator and a driver {[}5{]}. The
driver types at the computer while the navigator tells them what to
type, thus ensuring participation of both partners. Midway through the
activity, the presenter can have the partners switch roles.
Intentionally forming pairs can also engineer group interactions by
breaking up cliques, avoiding potentially disruptive partnerships, or
pairing reliable role models with new group members.

\subsection{Rule 9: Prepare in advance to maximize
participation}\label{rule-9-prepare-in-advance-to-maximize-participation}

It is not possible to fully participate in a Journal Club discussion
about a paper that the participant has not read. In that context, coming
to Code Club without having installed a dependency is similar to asking
a simple question about the Journal Club paper. Both instances show a
lack of respect and preparation. Just as a presenter must follow Rule 6
to provide materials ahead of time, participants must review the code in
advance, download the data sets, install the necessary packages, and
perhaps read up on the topic. If the Code Club is based on a paper or
chapters in a data science book (e.g., {[}6--8{]}) the participants
should read them before the session and consider how they might
incorporate the concepts into their own work.

\subsection{Rule 10: Participate}\label{rule-10-participate}

An essential ingredient of any Code Club is active participation from
all parties. Having an open laptop on the table and permission to use it
can feel like an invitation to get distracted by other work, emails, and
browsing the internet; fight that urge and focus on the presenter's
goals. Be respectful and allow your partner to contribute. Speak up for
yourself and force your partner to let you contribute. If the material
seems too advanced for you, it can be frustrating, and tempting to
mentally check out. Fortuntately, programming languages like R and
Python are generally expressive, which should allow you to engage with
the logic, even if the syntax is too advanced. Frankly, understanding
the logic of when to use one modeling approach over another is more
important than how to use it. If you understand the ``why'', the ``how''
will quickly follow. More experienced participants should aim to
communicate feedback and coding suggestions at a level that all
participants can understand and engage in. Regardless of skill levels,
your partner and the presenter put themselves at risk by revealing what
they do or do not know. Encourage them, and show your gratitude for
helping you learn something new, by fully participating in each Code
Club.

\subsection{Conclusion}\label{conclusion}

The most important rules are the first and last. Members of the Code
Club need to feel comfortable with other group members and sufficiently
empowered to try something new or ask for help. Aside from growing in
our programming skills, we have noticed two other benefits that help
create a positive work culture. First, we have intentionally interviewed
postdoc candidates on the days we hold Code Club. We make it clear that
they are not being assessed on their coding skills, but instead use it
as an opportunity to see how the candidate interacts with other members
of the research group in a relatively low stakes environment. At the
same time, the candidate can learn about the culture of the research
group through active participation. Second, members of other research
groups that feel isolated in growing their skills have integrated
themselves into our Code Club. This speaks to both the broader need for
Code Club and the likelihood of success when expanded to include a
larger group of individuals with broader research interests. Ultimately,
Code Club has improved the overall data analysis skills, community, and
research success of our lab by empowering researchers to seek help from
their colleagues.

\subsection{Acknowledgements}\label{acknowledgements}
